How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-05-01 tse-paper/content.tex
address comments, spellchecking

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
% Moritz: I think best to decided once we read the paper
% top to bottom.
%For now, it seems ok
% Needs to go Threats or intro
% TODO Caro: I moved it to the intro
% --- please check how it works for you
Retrospectively analyzing the build logs is not the only solution to
the problem---an alternate would be to integrate live plugins into the
build process
 such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin} or
Deflaker by Bell et al.~\cite{bell2018deflaker}.

Their advantage is that they can precisely report the information which
is required.
However, such plugins are not available for all build tools, they are
non-trivial to implement one-off solutions, and they
do not enable the analysis of historical builds.
When using such a plugin, one has to decide
beforehand which information should be reported, whereas
build log analysis enables us to leverage all information
present in the build logs.
Moreover, build log analyis is the more general soltuion.
While they serve an important niche, custom plugins do not make build
log analysis redundant.

Despite its central role in helping developers and in enabling
novel, onward processing applications,
the process of how to
% TODO
%automatically
% ^ we survey manual as well, though leaving it out includes too much
% systematically twice -- avoid this
analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-research to analyze build logs?
\end{simplebox}

@@

When analyzing system log files, the first goal is generally to
separate the constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
The output is an
ordered list of timed events and their corresponding parameter
values~\cite{he2016evaluation}.
This structured log can serve as
input to various machine learning and data mining processes.
Apart
from a large body of research on system logs, an equally mature amount
of commercial-grade and
opensource tooling exists to analyze system logs---such as, for
example, Elasticsearch, Logstash and Kiabana (the so-called ELK
stack)~\cite{sanjappa2017analysis,bajer2017building}.

@@

\Cref{fig:litsur:use} shows the aim for the
% TODO: isn't this discussion already?
% Moritz: Not sure what is meant by that?
investigated works to analyze build logs, and the result also explains
the prevalence of classification in \Cref{fig:litsur:kind}.
The majority of works (46\%) used the outcome of their build log
analysis to drive further research---and it is science's aim to
classify and abstract.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other works (10\%) used the gained insights to give feedback to
developers.
For 8\% of the works was the retrieved information the input
to a next automatic onward processing step:
works chose
tests that should be part of a reduced, but still effective test
suite~\cite{shi2018evaluating} or filled     a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

@@

The most mentioned technique (34\%) was using a parser.
In this tag, we also
included
rather imprecise statements like ``we parse the build
logs''~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they described
a complex parser (including a tokenizer, lexer, etc.) or refered to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``Regular Expression'' instead of the tag ``Parser.''
25\% of the works described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with a manual categorization of build logs and selected
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% of the investigated works
employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

Only 25\% of the works claimed their implementation is available.
Moreover, we
saw some reuse of the few available build log analysis tools.
% TODO The numbers here might look a bit strange
% --- we say five but really it is six works we reference.
% I assume this is because of only including the oldest ref?
% Still, I think we need to fix the inconsistency.
Five works employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017travistorrent,beller2017oops,
orellana2017differences,zhao2018comparing} or
an updated version of them~\cite{rott2019empirische,
shi2018evaluating}, two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise},
and
another two
\emph{MAKAO}~\cite{wen2018blimp,adams2007design,adams2007makao}.

@@

In 38\% of cases, the works collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the works analyzed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the works
and we were not always able to
extract it confidently.
23 of the works analyzed more than a thousand build logs, while one
investigated up to 122 million logs.
Of the surveyed works, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
% TODO	include the several others as references?
and several others have studied.

@@

% TODO passive ...
The literature survey shows the need for
build log
analysis techniques that are \emph{automated},
\emph{build tool-agnostic}
and can be \emph{configured with little effort}.

@@

In this section, we formally define the concept of chunk
retrievals from build logs.
Then, we motivate and present the choice of three such techqniues that
we implemented for this study in detail.

% TODO Caro: Not sure what to do with this ...
%% In this section,
%% we propose three so-called \emph{chunk retrieval techniques},
%% which retrieve a specific substring (chunk)
%% from a build log
%% % TODO passive ....
%% The techniques we propose run automatically
%% and can be easily configured through examples.
%% % TODO this should go under characteristics and not be in an introduction
%% to the section ...
%% An example in this case consists of a build log together with the
%% chunk that should be retrieved.
%% A user---a developer or a researcher who wants to analyze
%% build logs---provides these examples, which are easy to
%% adapt whenever the analysis technique needs
%% to support new cases.

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
% TODO (resolved) Never define a concept with the
% words used in the term
% Use a thesaurus to find synonyms

@@

Techniques based on custom parsers (\#1
in \Cref{tab:litsur:techniques}) are inadquate for our comparison,
since their creation is manual and such a technique would not
generalize to the wide varitey of build logs we are interested in.
The
closest we could come to the second entry, regular expression, is a
technique to automatically derive them: in particular, we implement a
technique based on
\emph{Program Synthesis by Example (PBE)}.
This mimics regular expressions, while simplifying their development.

@@

\begin{table*}[htb]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}llXll@{}}
\toprule
Name & Acronym & Identification Technique & Granularity &
Configuration \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
% TODO: TF-IDF still not defined
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:techniques}
\end{table*}

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
algorithm weighs the vectors using \emph{term frequency-inverse
document frequency}, a best practice for natural
language queries~\cite{lee1997document}.

@@

For each build log, we manually marked
the chunk that describes why the build
% TODO add back ref
failed---the information targeted by most of the works in our
literature survey.
To be able to configure the keyword search technique,
the data set associates each chunk with keywords that we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

>>>>

Retrospectively analyzing the build logs is not the only solution to
the problem---an alternate would be to integrate live plugins into the
build process
 such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin} or
Deflaker by Bell et al.~\cite{bell2018deflaker}.

Their advantage is that they can precisely report the information which
is required.
However, such plugins are not available for all build tools, they are
non-trivial to implement one-off solutions, and they
do not enable the analysis of historical builds.
When using such a plugin, one has to decide
beforehand which information should be reported, whereas
build log analysis enables us to leverage all information
present in the build logs.
Moreover, build log analysis is the more general solution.
While they serve an important niche, custom plugins do not make build
log analysis redundant.

Despite its central role in helping developers and in enabling
novel, onward processing applications,
the process of how to
automatically
analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-research to analyze build logs?
\end{simplebox}

@@

When analyzing system log files, the first goal is generally to
separate the constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
The output is an
ordered list of timed events and their corresponding parameter
values~\cite{he2016evaluation}.
This structured log can serve as
input to various machine learning and data mining processes.
Apart
from a large body of research on system logs, an equally mature amount
of commercial-grade and
open source tooling exists to analyze system logs---such as, for
example, Elasticsearch, Logstash and Kiabana (the so-called ELK
stack)~\cite{sanjappa2017analysis,bajer2017building}.

@@

\Cref{fig:litsur:use} shows the aim for the
investigated works to analyze build logs, and the result also explains
the prevalence of classification in \Cref{fig:litsur:kind}.
The majority of works (46\%) used the outcome of their build log
analysis to drive further research---and it is science's aim to
classify and abstract.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other works (10\%) used the gained insights to give feedback to
developers.
For 8\% of the works was the retrieved information the input
to a next automatic onward processing step:
works chose
tests that should be part of a reduced, but still effective test
suite~\cite{shi2018evaluating} or filled     a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

@@

The most mentioned technique (34\%) was using a parser.
In this tag, we also
included
rather imprecise statements like ``we parse the build
logs''~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they described
a complex parser (including a tokenizer, lexer, etc.) or referred to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``Regular Expression'' instead of the tag ``Parser.''
25\% of the works described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with a manual categorization of build logs and selected
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% of the investigated works
employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

Only 25\% of the works claimed their implementation is available.
Moreover, we
saw some reuse of the few available build log analysis tools.
Five works employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017oops,
orellana2017differences,zhao2018comparing} or
an updated version of them~\cite{rott2019empirische,
shi2018evaluating}, two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise},
and
another two
\emph{MAKAO}~\cite{wen2018blimp,adams2007design}.

@@

In 38\% of cases, the works collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the works analyzed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the works
and we were not always able to
extract it confidently.
23 of the works analyzed more than a thousand build logs, while one
investigated up to 122 million logs.
Of the surveyed works, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
and several others have studied~\cite{hassan2017change,
hassan2019tackling,hassan2017automatic,orellana2017differences,
haghighatkhah2018test,madeyski2017continuous,ghaleb2019empirical,
santolucito2018statically,gallaba2018noise,shi2018evaluating}.

@@

The literature survey shows the need for
build log
analysis techniques that are \emph{automated},
\emph{build tool-agnostic}
and that a user can \emph{configure with little effort}.

@@

In this section, we formally define the concept of chunk
retrievals from build logs.
Then, we motivate and present the choice of three such techniques that
we implemented for this study in detail.

% TOO Caro: Not sure what to do with this ...
%% In this section,
%% we propose three so-called \emph{chunk retrieval techniques},
%% which retrieve a specific substring (chunk)
%% from a build log
%% % TOO passive ....
%% The techniques we propose run automatically
%% and can be easily configured through examples.
%% % TOO this should go under characteristics and not be in an
% introduction
%% to the section ...
%% An example in this case consists of a build log together with the
%% chunk that should be retrieved.
%% A user---a developer or a researcher who wants to analyze
%% build logs---provides these examples, which are easy to
%% adapt whenever the analysis technique needs
%% to support new cases.

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}

@@

Techniques based on custom parsers (\#1
in \Cref{tab:litsur:techniques}) are inadequate for our comparison,
since their creation is manual and such a technique would not
generalize to the wide variety of build logs we are interested in.
The
closest we could come to the second entry, regular expression, is a
technique to automatically derive them: in particular, we implement a
technique based on
\emph{Program Synthesis by Example (PBE)}.
This mimics regular expressions, while simplifying their development.

@@

\begin{table*}[htb]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}llXll@{}}
\toprule
Name & Acronym & Identification Technique & Granularity &
Configuration \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:techniques}
\end{table*}

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
algorithm weighs the vectors using \emph{term frequency-inverse
document frequency (TF-IDF)}, a best practice for natural
language queries~\cite{lee1997document}.

@@

For each build log, we manually marked
the chunk that describes why the build
failed---the information targeted by most of the works in our
literature survey (see \Cref{sec:lit-sur:discussion}).
To be able to configure the keyword search technique,
the data set associates each chunk with keywords that we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.
